---
title: "Discover ERPscope - plot_erp, plot_difference & generate_stats_table"
author: "Alexandre Herbay"
date: "23/05/2025"
output:
  pdf_document:
    toc: yes
    toc_depth: '2'
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 2
    number_sections: yes
    code_folding: hide
    theme: united
    fig_width: 8
    fig_height: 6
editor_options:
  chunk_output_type: inline
---

# Setup

Run this code manually just once if the package is not already installed

```{r, eval=FALSE}
  detach("package:erpscope", unload=TRUE) 
  devtools::install_github("aherbay/erpscope", ref= "workshop_2025")
  library(erpscope)
```

```{r, eval=FALSE}
  install.packages('gifski')
```


```{r setup, include=FALSE}
library(erpscope)
```

Let's discover the main capabilities

In this part, we will use the relpriming dataset included in the package. This dataset included the data from `r length(unique(relpriming$Subject))` participants.
The preprocessing baseline was [-500ms,-300ms]. 

# Basic plot in the window

We minimally need to indicate 3 things : which data we use, the condition to plot and the electrode to plot

## Single electrode

```{r}
head(relpriming, 10)
```


```{r single_electrode}
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = Cz
  )
```

## Layout of 9 electrodes 
```{r layout_9elec }
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = lay_9el_FCP_3z4
  )

```

## Layout of 64 electrodes 

```{r layout_64elec , eval=FALSE}
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = lay_64e_waveguard_cap,
             electrode_labels_font_size = 8
  )
```

## Determine which polarity is plotted up

```{r}
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = Cz,
             polarity_up = "positive"
  )
```


# Explore file options

## Basic plot in a file

possible file format:  pdf, jpeg, svg, tiff, png, bmp, eps/ps, tex (pictex)

```{r}
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = lay_9el_FCP_3z4, 
           output_format = 'pdf'
)
```

## File name

By default it is automatic with the following format:
time - plot type - dataset - condition - electrode layout

```{r, eval=FALSE}
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = lay_9el_FCP_3z4, 
             output_format = 'pdf',
             output_file_name = "a_much_better_title3"
  )
```

## Image output

```{r, eval=FALSE}
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = lay_9el_FCP_3z4, 
            output_format = 'jpeg',
  )
```

## JPEG output

```{r, eval=FALSE}
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = lay_9el_FCP_3z4, 
             plot_subtitle = 'my_subtitle',
             plot_subtitle_font_color = 'black',
  )
```


## File dimension

by default : 8.5 by 11 inches

```{r}
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = lay_9el_FCP_3z4, 
             output_format = 'pdf',
             plot_height = 8.5,
             plot_width = 20 
  )
```

## Create a file in a subfolder

possible file format:  pdf, jpeg, svg, tiff, png, bmp, eps/ps, tex (pictex)

```{r}
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = lay_9el_FCP_3z4, 
           output_format = 'pdf',
           output_subfolder = "plots/"
)
```


# Explore graphic options

## line colors
```{r}
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           line_colors = c('black','green','blue')
  
)
```


## line types

one single value: same type for all lines, 
a vector: different values for each level of the condition to plot (in the same order as levels of the condition)

https://r-charts.com/base-r/line-types/

c('solid', 'dotted' , 'dashed','longdash','F1')
```{r}
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           line_type = c('solid','dashed','1342')
  
)
```

## line thickness

line_thickness argument
one single value: same type for all lines, 
a vector: different values for each level of the condition to plot (in the same order as levels of the condition)

```{r}
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           line_thickness = 1.2
  
)

plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           line_thickness = c(0.5, 0.75,1.2)
  
)

```

## plot background

default: "grid",  alternative : "dark" or "white"

```{r}
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = Cz,
             plot_background =  "dark"
  )
  
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = Cz,
             plot_background =  "white"
  )
```

# Explore axis options

## voltage axis

voltage_axis_title : default Voltage amplitude (microV)
voltage_axis_tick_mark_labels_font_size
voltage_scale_limits

```{r}

  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = Cz,
             voltage_axis_title =  "Voltage",
             voltage_scale_limits = c(10,-10)
  )

```

            
## time axis

time_axis_title

time_scale_limits
time_labels

time_axis_tick_mark_labels_font_size

```{r}
 plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = Cz,
             time_axis_title =  "Temps (en ms)",
             time_scale_limits = c(0,700),
             time_labels = c(0,200,300,500,700)
  )

```

## axis font sizes

```{r}
 plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = lay_9el_FCP_3z4,
              time_axis_tick_mark_labels_font_size = 9,
             voltage_axis_tick_mark_labels_font_size = 9
  )

```


# Explore text options


## title and subtitle

title by default: automatic "ERP: ", vary,"by",condition_to_plot 
subtitle by default: Date - Baseline:[beginning ms;end ms] - dataset: dataset_name with X subjects

you can also change: plot_title_font_size, plot_title_font_color, plot_subtitle_font_size, plot_subtitle_font_color

```{r}
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           plot_title = "My own title",
           plot_subtitle = "My own subtitle"
  
)
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           plot_title = "",
           plot_subtitle = ""
  
)
```



## change overall font size

```{r}
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           text_font_size = 16
  
)

```

## electrode labels

```{r}
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           electrode_labels_font_size = 16,
           electrode_labels_font_color = 'grey'
  
)

```

# Explore custom labels options

 custom_labels list of custom label list. Each custom labels should have the structure: list(start_time, end_time, "label")
 custom_labels_vertical_position 'auto_top', 'auto_bottom' or custom position for the center of the label (in microVolts)
 custom_labels_height 'auto' or custom height (in microVolts)
 custom_labels_fill_color 'grey'
 custom_labels_font_size 1.9
 custom_labels_line_style "dashed" # not implemented
 custom_labels_line_color "black" # not implemented

```{r}
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           polarity_up =  'negative',
           custom_labels = list(list(0,200,"Target")),
           custom_labels_vertical_position = 'auto_bottom',
           custom_labels_height = 'auto'
)

```


# Explore variability information options

## mean_sdl

```{r}

plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           variability_info = "mean_sdl"
) 

```

'mean_cl_boot', "mean_cl_normal"

## mean_cl_normal

```{r}
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = Cz,
             variability_info = "mean_cl_normal"
  )
```

## mean_cl_boot

```{r}
  plot_erp ( data = relpriming,
             condition_to_plot = "Pair.Type",
             electrodes_layout = Cz,
             variability_info = "mean_cl_boot"
  )
```

## within_subject_ci

```{r}
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           variability_info = "within_subject_ci"
)
```

# Explore display of info about preprocessing baseline

 prepro_bsl_display boolean, defines whether an indication of preprocessing baseline is displayed or not
 prepro_bsl_time_window vector defining the time limits of the preprocessing baseline (e.g.,c(-500, -300))
 prepro_bsl_fill_color = "#DEE5ED"
 prepro_bsl_label_fill_alpha value between 0 and 1 defining the transparency of the preprocessing baseline indication
 prepro_bsl_label  ""
 prepro_bsl_label_font_size  2
 prepro_bsl_label_text_color  "blue"
 prepro_bsl_label_vertical_position vertical position of the center of the rectangle
 prepro_bsl_vertical_limits  c(-0.5,0.5)

```{r}

plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           prepro_bsl_display = T,
           prepro_bsl_time_window = c(-500,-300),
           prepro_bsl_vertical_limits = c(6.5,7),
           prepro_bsl_fill_color = 'grey'
           
)
```

# Explore simulated baseline options


## simulate a new baseline

  simul_bsl_active = FALSE,
  simul_bsl_time_window = c(-200, 0),
  simul_bsl_label_fill_color = "#8282DF",
  simul_bsl_label_fill_alpha = 0.9,
  simul_bsl_label_text = "",
  simul_bsl_label_font_size = 4,

```{r}

plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz,
           simul_bsl_active = T,
           simul_bsl_time_window = c(-0,200),
           simul_bsl_vertical_limits = c(4.5,5),
           simul_bsl_label_fill_color = 'grey',
           simul_bsl_label_text = 'simulated baseline',
           simul_bsl_label_font_size = 1.7
           
)
```


## Compare simulated and preprocessing baseline

install.packages('gifski')

```{r,  animation.hook="gifski", eval=FALSE}
 
plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz ,
           prepro_bsl_display = T,
           prepro_bsl_time_window = c(-500,-300),
           prepro_bsl_vertical_limits = c(6.5,7),
           voltage_scale_limits = c(7,-7)
)

plot_erp ( data = relpriming,
           condition_to_plot = "Pair.Type",
           electrodes_layout = Cz ,
           simul_bsl_active = TRUE,
           simul_bsl_time_window = c(-200, 0),
           simul_bsl_vertical_limits = c(6.5,7),
           voltage_scale_limits = c(7,-7)
)


```


# Explore plot_difference

## Some electrodes

```{r}
plot_difference( data = relpriming,
                 conditionToPlot = Pair.Type,
                 levelA = Unrelated ,
                 levelB = Consistent,
                 output_type ='pdf',
                 vary= Voltage,
                 group_var = Subject,
                 show_group_obs = FALSE ,
                 labels_vertical_position = 'auto',
                 labels_height = 'auto',
                 baseline= c(-500,-200),
                 time_labels_interval = 200,
                 custom_labels = list(list(-450,-250,"Prime"),list(0,200,"Target")),
                 electrodes_to_display = c("F3", "Fz", "F4","C3", "Cz","C4", "P3", "Pz", "P4"),
                 plotname = 'auto'
) 
```

# ROI

```{r}
plot_difference( data = relpriming,
                 conditionToPlot = Pair.Type,
                 ant_levels= Anteriority.Levels,
                 med_levels= Mediality.Levels,
                 levelA = Unrelated ,
                 levelB = Consistent,
                 output_type ='pdf',
                 vary= Voltage,
                 group_var = Subject,
                 show_group_obs = FALSE ,
                 labels_vertical_position = 'auto',
                 labels_height = 'auto',
                 baseline= c(-500,-200),
                 time_labels_interval = 200,
                 custom_labels = list(list(-450,-250,"Prime"),list(0,200,"Target")),
                 plotname = 'auto'
) 
```

## with subject average


```{r}
plot_difference( data = relpriming,
                 conditionToPlot = Pair.Type,
                 levelA = Unrelated ,
                 levelB = Consistent,
                 output_type ='pdf',
                 vary= Voltage,
                 group_var = Subject,
                 show_group_obs = TRUE ,
                 labels_vertical_position = 'auto',
                 labels_height = 'auto',
                 baseline= c(-500,-200),
                 topoplots_time_windows = list(c(-250,-150),c(-150,50),c(50,200),c(200,300),c(300,500),c(500,700),c(700,900)),
                 topoplots_scale = c(-2,2),
                 time_labels_interval = 200,
                 custom_labels = list(list(-450,-250,"Prime"),list(0,200,"Target")),
                 electrodes_to_display = c("F3", "Fz", "F4","C3", "Cz","C4", "P3", "Pz", "P4"),
                 plotname = 'auto'
) 

```

## with t-test 

```{r}
plot_difference( data = relpriming,
                 conditionToPlot = Pair.Type,
                 levelA = Unrelated ,
                 levelB = Consistent,
                 output_type ='pdf',
                 ant_levels= Anteriority.Levels,
                 med_levels= Mediality.Levels,
                 vary= Voltage,
                 group_var = Subject,
                 show_group_obs = TRUE ,
                 labels_vertical_position = 'auto',
                 labels_height = 'auto',
                 baseline= c(-500,-200),
                 topoplots_time_windows = list(c(-250,-150),c(-150,50),c(50,200),c(200,300),c(300,500),c(500,700),c(700,900)),
                 topoplots_scale = c(-2,2),
                 time_labels_interval = 200,
                 custom_labels = list(list(-450,-250,"Prime"),list(0,200,"Target")),
                 electrodes_to_display = c("F3", "Fz", "F4","C3", "Cz","C4", "P3", "Pz", "P4"),
                 plotname = 'auto',
                 show_t_test = T,
                 t_test_threshold = 0.01
) 

```

# plot difference maps

```{r}
plot_difference_maps(  data = relpriming,
                         conditionToPlot = Pair.Type,
                         levelA = Unrelated ,
                         levelB = Consistent,
                         output_type ='window',
                         fixed= c(-300,900,100), # init_time, end_time, step
                         topoplots_scale = c(-2,2),
                         plotname = 'auto'
)

```


# generate stats table

```{r}
 generate_ERP_stats_table( 
      data = relpriming,
      model_structure = "Voltage ~   Pair.Type + (1+ Pair.Type|Subject)",
      timeWindowMode="custom",
      custom_TW =  list(c(-300,-150),c(-150,50),c(50,200),c(200,300),c(300,500),c(500,700)),
      output_name="2020_07_02_PairTypeModels.html"
 ) 
```



